---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

This document contains the statistical tests and results for alpha diversity measures associated with POC_2016.
By Rebecca Maher


```{r}
library('lme4')
library('emmeans')
library('lmerTest')
library('dplyr')
```
## Load the alpha diversity datafile.
```{r}
alphadiv <- read.csv(file = "~/data/alphadiv.csv")
summary(alphadiv)
```

## Statistical analysis of full data set with growth rate and symbiodinium density.
### Growth rates
####First test the effect of growth rates on chao1
```{r}
growth_chao1 <- lmer(logrich ~ growth + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(growth_chao1)
summary(growth_chao1)

growth_chao1.2 <- lm(logrich ~ growth, data = alphadiv)
anova(growth_chao1.2)
summary(growth_chao1.2)

# Just checking if I'm using the right response ~ predictor structure
growth_chao1.3 <- lm(growth ~ logrich, data = alphadiv)
anova(growth_chao1.3)
summary(growth_chao1.3)
```
These results show that growth rates are not a predictor of richness (log-transformed chao1 index values).

####Next, test the effect of growth rates on Simpson's diversity
```{r}
growth_simp <- lmer(asineven ~ growth + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(growth_simp)
summary(growth_simp)

growth_simp1 <- lm(asineven ~ growth, data = alphadiv)
anova(growth_chao1)
summary(growth_chao1)
```
These results show that growth rates are not a predictor of simpson's diversity (asin-transformed).

####Lastly, test the effect of growth rates on Faith's phylogenetic diversity
```{r}
growth_faith <- lmer(logfaithPD ~ growth + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(growth_faith)
summary(growth_faith)

growth_faith1 <- lm(logfaithPD ~ growth, data = alphadiv)
anova(growth_faith1)
summary(growth_faith1)
```
These results show that growth rates are not a predictor of phylogenetic diversity (log-transformed faith PD values).


### Symbiodinium densities
####First test the effect of symbiodinium densities on chao1
```{r}
zcount_chao1 <- lmer(logrich ~ zcount + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(zcount_chao1)
summary(zcount_chao1)

zcount_chao1.2 <- lm(logrich ~ zcount, data = alphadiv)
anova(zcount_chao1.2)
summary(zcount_chao1.2)

# Just checking if I'm using the right response ~ predictor structure
zcount_chao1.3 <- lm(zcount ~ logrich, data = alphadiv)
anova(zcount_chao1.3)
summary(zcount_chao1.3)
```
These results show that symbiodinium densities are not a predictor of richness (log-transformed chao1 index values).

####Next, test the effect of symbiodinium densities on Simpson's diversity
```{r}
zcount_simp <- lmer(asineven ~ zcount + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(zcount_simp)
summary(zcount_simp)

zcount_simp1 <- lm(asineven ~ zcount, data = alphadiv)
anova(zcount_chao1)
summary(zcount_chao1)
```
These results show that symbiodinium densities are not a predictor of simpson's diversity (asin-transformed).

####Lastly, test the effect of symbiodinium densities on Faith's phylogenetic diversity
```{r}
zcount_faith <- lmer(logfaithPD ~ zcount + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(zcount_faith)
summary(zcount_faith)

zcount_faith1 <- lm(logfaithPD ~ zcount, data = alphadiv)
anova(zcount_faith1)
summary(zcount_faith1)
```
These results show that symbiodinium densities are not a predictor of phylogenetic diversity (log-transformed faith PD values).


### Healing rates
####Subset data to only include samples with a healing rate (scarred samples)
```{r}
alphadiv <- na.omit(alphadiv)
summary(alphadiv)
```
####First test the effect of healing rates on chao1
```{r}
healing_chao1 <- lmer(logrich ~ healing + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(healing_chao1)
summary(healing_chao1)

healing_chao1.2 <- lm(logrich ~ healing, data = alphadiv)
anova(healing_chao1.2)
summary(healing_chao1.2)

# Just checking if I'm using the right response ~ predictor structure
healing_chao1.3 <- lm(healing ~ logrich, data = alphadiv)
anova(healing_chao1.3)
summary(healing_chao1.3)
```
These results show that healing rates are not a predictor of richness (log-transformed chao1 index values).

####Next, test the effect of healing rates on Simpson's diversity
```{r}
healing_simp <- lmer(asineven ~ healing + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(healing_simp)
summary(healing_simp)

healing_simp1 <- lm(asineven ~ healing, data = alphadiv)
anova(healing_chao1)
summary(healing_chao1)
```
These results show that healing rates are not a predictor of simpson's diversity (asin-transformed).

####Lastly, test the effect of healing rates on Faith's phylogenetic diversity
```{r}
healing_faith <- lmer(logfaithPD ~ healing + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(healing_faith)
summary(healing_faith)

healing_faith1 <- lm(logfaithPD ~ healing, data = alphadiv)
anova(healing_faith1)
summary(healing_faith1)
```
These results show that healing rates are not a predictor of phylogenetic diversity (log-transformed faith PD values).

#### Now I'm going to look at treatment effects on alpha diversity of healing rate data.
Still using the subsetted alphadiv dataframe to only include samples with a healing rate.
```{r}
# Chao1
healing_trt_c <- lmer(logrich ~ temp*nutrient + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(healing_trt_c)
summary(healing_trt_c)

healing_trt_c1 <- lm(logrich ~ temp*nutrient, data = alphadiv)
anova(healing_trt_c1)
summary(healing_trt_c1)

# Simpson's diversity
healing_trt_s <- lmer(asineven ~ temp*nutrient + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(healing_trt_s)
summary(healing_trt_s)

healing_trt_s1 <- lm(asineven ~ temp*nutrient, data = alphadiv)
anova(healing_trt_s1)
summary(healing_trt_s1)

# Faith's phylogenetic diversity
healing_trt_f <- lmer(faithPD ~ temp*nutrient + (1|tank) + (1|colony), data = alphadiv, REML = T)
anova(healing_trt_f)
summary(healing_trt_f)

healing_trt_f1 <- lm(faithPD ~ temp*nutrient, data = alphadiv)
anova(healing_trt_f1)
summary(healing_trt_f1)
```
Nothing was significant here.

Now trying with pairwise t.tests
```{r}
pairwise.t.test(alphadiv$logrich, alphadiv$nutrient, p.adjust.method = "fdr")
pairwise.t.test(alphadiv$asineven, alphadiv$nutrient, p.adjust.method = "fdr")
pairwise.t.test(alphadiv$logfaithPD, alphadiv$nutrient, p.adjust.method = "fdr")

pairwise.t.test(alphadiv$logrich, alphadiv$temp, p.adjust.method = "fdr")
pairwise.t.test(alphadiv$asineven, alphadiv$temp, p.adjust.method = "fdr")
pairwise.t.test(alphadiv$logfaithPD, alphadiv$temp, p.adjust.method = "fdr")
```
Also no other significant results.

#### Now going to calculate summary statistics for samples with a healing rate
```{r}
sderr <- function(x) {sd(x)/sqrt(length(x))}

alphadiv_grp <- group_by(alphadiv, nutrient)

mean_stderr_grp_c <- function(subdf) {
  meanm <- mean(subdf$richness)
  stderr <- sderr(subdf$richness)
  ret_df <- data.frame(mean = meanm, se = stderr)
  return(ret_df)
}

chao_stats <- do(alphadiv_grp, mean_stderr_grp_c(.))

mean_stderr_grp_s <- function(subdf) {
  meanm <- mean(subdf$evenness)
  stderr <- sderr(subdf$evenness)
  ret_df <- data.frame(mean = meanm, se = stderr)
  return(ret_df)
}

simp_stats <- do(alphadiv_grp, mean_stderr_grp_s(.))

mean_stderr_grp_f <- function(subdf) {
  meanm <- mean(subdf$faithPD)
  stderr <- sderr(subdf$faithPD)
  ret_df <- data.frame(mean = meanm, se = stderr)
  return(ret_df)
}

faith_stats <- do(alphadiv_grp, mean_stderr_grp_f(.))
chao_stats
simp_stats
faith_stats

```



